# Clustering
http://www.sthda.com/english/articles/25-cluster-analysis-in-r-practical-guide/111-types-of-clustering-methods-overview-and-quick-start-r-code/

```{r}
library(ggplot2)
library("cluster")
library("factoextra")
library("magrittr")
```

```{r}
data("USArrests")
my_data <- USArrests
head(my_data,10)
```

```{r}
my_data <- na.omit(my_data)
my_data <- scale(my_data)
head(my_data)
```
## Distance Computation

#### Correlation based distance for numeric values
```{r}
sub_data <- my_data[1:10,]
res.dist <- get_dist(sub_data, stand = TRUE, method = "pearson")
fviz_dist(res.dist)
```
> Red: high similarity (ie: low dissimilarity) | Blue: low similarity

## Partitioning clustering
Partitioning clustering are clustering methods used to classify observations, 
within a data set, into multiple groups based on their similarity

### K Means Clustering:
http://www.sthda.com/english/articles/27-partitioning-clustering-essentials/87-k-means-clustering-essentials/

#### Determining the ultimate number of clusters
```{r}
fviz_nbclust(my_data, kmeans, method = "wss")
```

```{r}
set.seed(123)
km.res <- kmeans(my_data, 4, nstart = 25)
km.res
```

```{r}
# Visualize
fviz_cluster(km.res, data = my_data,
             ellipse.type = "convex",
             palette = "jco",
             ggtheme = theme_minimal())

fviz_cluster(km.res, data = my_data,
             palette = c("#2E9FDF", "#00AFBB", "#E7B800", "#FC4E07"), ellipse.type = "euclid", # Concentration ellipse
             #star.plot = TRUE, # Add segments from centroids to items repel = TRUE, # Avoid label overplotting (slow)
             ggtheme = theme_minimal()
)
```

### PAM Clustering:
http://www.sthda.com/english/articles/27-partitioning-clustering-essentials/88-k-medoids-essentials/

#### Estimating the optimal number of clusters
```{r}
fviz_nbclust(my_data, pam, method = "silhouette")+
theme_classic()
```


```{r}
# Compute PAM
pam.res <- pam(my_data, 2)
pam.res
```


```{r}
# Visualize
fviz_cluster(pam.res)
```




